---
title: "Nonlinear"
author: "JunLu"
date: "4/1/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE,
                      message = FALSE
                      )
library(caret)
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .2) 
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1) 
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)

library(tidyverse)
library(caret)
library(splines)
library(gam)
library(mgcv)
library(boot)
library(pdp)
library(earth)
```


## Import the train dataset

```{r}
train = read_csv("./non-linear/train.csv")

y = train$transformed_value
options(na.action = 'na.pass')
x = model.matrix(transformed_value ~ ., train)[,-1] 
```

## Set a random seed

```{r}
set.seed(2)
```


## Generallized additive model (GAM)

### Use caret package
```{r, eval=FALSE}
ctrl1 <- trainControl(method = "cv", number = 10)
gam.fit = train(x, y,
                preProcess = "medianImpute",
                method = "gam",
                tuneGrid = data.frame(method = "GCV.Cp", 
                                      select = c(TRUE,FALSE)), 
                trControl = ctrl1)

save(gam.fit, file = "./gam_fit.rda")
```

```{r}
load(file = "./gam_fit.rda")
gam.fit$bestTune
gam.fit$finalModel
```


```{r}
par(mfrow = c(1,3)) 
plot(gam.fit$finalModel)
```


## Multivariate Adaptive Regression Splines (MARS)

```{r, eval=FALSE}
ctrl1 <- trainControl(method = "cv", number = 10)
mars_grid = expand.grid(degree = 1:2,
                        nprune = 2:38)
mars.fit = train(x, y,
                 method = "earth",
                 preProcess = "medianImpute",
                 tuneGrid = mars_grid,
                 trControl = ctrl1
                 )
save(mars.fit, file = "./earth.rda")
```

```{r}
load(file = "./earth.rda")
summary(mars.fit)
mars.fit$bestTune
```

```{r}
p1 = partial(mars.fit, pred.var = c("age"), grid.resolution = 10) %>% autoplot()
p2 = partial(mars.fit, pred.var = c("potential"), grid.resolution = 10) %>% autoplot()
p3 = partial(mars.fit, pred.var =c("agility"), grid.resolution = 10) %>% autoplot()
p4 = partial(mars.fit, pred.var =c("balance"), grid.resolution = 10) %>% autoplot()
p5 = partial(mars.fit, pred.var =c("gk_diving"), grid.resolution = 10) %>% autoplot()
p6 = partial(mars.fit, pred.var =c("gk_handling"), grid.resolution = 10) %>% autoplot()
p7 = partial(mars.fit, pred.var =c("gk_kicking"), grid.resolution = 10) %>% autoplot()
p8 = partial(mars.fit, pred.var =c("gk_positioning"), grid.resolution = 10) %>% autoplot()
p9 = partial(mars.fit, pred.var =c("gk_reflexes"), grid.resolution = 10) %>% autoplot()
p10 = partial(mars.fit, pred.var =c("reactions"), grid.resolution = 10) %>% autoplot()
p11 = partial(mars.fit, pred.var =c("strength"), grid.resolution = 10) %>% autoplot()

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol = 3, nrow = 4)
```


## Compare those models
```{r}
bwplot(resamples(list(mars = mars.fit,gam = gam.fit)), 
       metric = "RMSE")
```


## Importance
```{r}
varImp(mars.fit)
varImp(gam.fit)
```


## Questions

1. As we can't use the test dataset to choose our final model, do we need to calculate test error for each model or just our final model. 

Do we write (linear -> nonlinear, if we know it is nonlinear why we try linear firstly)

depend on oursevles


2. How to know our model is good or not, just with a train error and a test error. (close and both small -> good, train << test overfitted, both large underfitted) However, how do we know it is small or large (MSE also depends on y)

can't

3. Which variables are important? Different models have different important variables. Do we only use the result of our final model?

4. how do we present our model in the report (coefficients? tunning parameter? plot?)


5. Model assumptions and limitations

*linear*
Multi linear: model linear in parameter (error term mean 0 constant variance uncorrelated)

lasso and ridge: linear (Anything else? Multicollinearity?)

pcr: linear (Anything else? Multicollinearity?)


*non-linear*
gam: Nonlinear relationship (Anything else?)
mars: Nonlinear relationship (Anything else)

6. How to choose tunning parameters?
like we try lambda from (e^-10, e^10), then we choose the best lambda by 10-fold cross vaildation.

7.How do we know model is enough flexible? How do we make prediction?


