---
title: "EDA"
date: "March 27, 2019"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE,
                      message = FALSE
                      )
```

#library packages

```{r}
library(caret)
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .2) 
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1) 
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)

library(tidyverse)
library(patchwork)
```


#Data cleaning

```{r}
fifa = read_csv("..\\Data\\CompleteDataset.csv")
```


```{r}
calc_expression = function(fmula) {
    eval(parse(text = fmula))
}

data = fifa %>% 
    janitor::clean_names() %>% 
    select(-c(x1, name, photo, flag, club_logo, wage, overall, id)) %>% 
    mutate(value = str_replace(value, "K", "/ 1000"),
           value = str_replace(value, "M", ""),
           value = str_replace(value, "€", "")) %>% 
    mutate(value = map(value, calc_expression),
           value = as.numeric(value)) %>% 
    mutate(value = readr::parse_number(value)) %>% 
    filter(preferred_positions == "GK") %>% 
    select(-(cam:st))

trans2int_cols = c(7:40)
trans2fct_cols = c(2, 4)
data[trans2int_cols] = map(data[trans2int_cols], as.integer)
data[trans2fct_cols] = map(data[trans2fct_cols], as.factor)

data = data %>% 
    mutate(nationality = as.character(nationality))

#combining different nations on the same continent into a new variable 
#which has fewer categories

eu = c("Germany", "Spain", "Italy", "Belgium", "Slovenia", "France", 
       "Czech Republic", "Croatia", "Switzerland", "Portugal", 
       "Denmark", "Poland", "Greece", "Bosnia Herzegovina", "England",
       "Norway", "Netherlands", "Finland", "Russia", "Turkey", "Ukraine",
       "Romania", "Albania", "Hungary", "Lithuania", 
       "Republic of Ireland", "Austria", "Sweden", "Wales", "Scotland",
       "Bulgaria", "Serbia", "Georgia", "Kosovo", "Slovakia", "Latvia",
       "Belarus", "FYR Macedonia", "Northern Ireland", "Iceland",
       "Luxembourg", "Montenegro", "Israel", "San Marino")
as = c("China PR", "Korea Republic", "Japan", "Oman", "Saudi Arabia",
       "Egypt", "Iran","Philippines", "India", "Lebanon", "Senegal",
       "Morocco",  "Comoros", "Nigeria", "Algeria", "Ivory Coast", 
       "Ghana", "DR Congo", "Benin", "Kenya", "Equatorial Guinea",
       "Gabon", " Burkina Faso", "Congo", "Tunisia", "Cape Verde", "Angola")
af = c("Cameroon", "South Africa", "Mali")
na = c("United States", "Guatemala", "Canada", "Puerto Rico",
       "Haiti", "Bermuda")
sa = c("Costa Rica", "Argentina", "Brazil", "Uruguay", "Chile", 
       "Colombia", "Mexico", "Venezuela", "Curacao", "Peru", 
       "Paraguay", "Ecuador", "Bolivia")
oc = c("Australia", "New Zealand")

nation_eu = function(name){
  if (name %in% eu)
      name_new = "eu"
  else if (name %in% as)
      name_new = "as"
  else if (name %in% na)
      name_new = "na"
  else if (name %in% sa)
      name_new = "sa"
  else if (name %in% oc)
      name_new = "oc"
  else name_new = "af"
  name_new
}


data = data %>% 
    mutate(nationality = map(.x = nationality, ~nation_eu(.x))) %>% 
    mutate(nationality = as.factor(unlist(nationality)))

    
```

#Split the data set into training and testing data

```{r}
set.seed(2)
trRows = createDataPartition(data$value, p = .75, list = FALSE)
data_split = data %>% 
    select(-club)
train = data_split[trRows,] 
test = data_split[-trRows,]
```

#Variable transformation

##Transform the response(value), based on the distribution of the response in the training dataset

```{r}
#The effect of log transformation is not ideal
#data = data %>%
#   mutate(value = log(value + 2)

p1 = train %>% 
    ggplot(aes(x = value)) + geom_density(fill = "navy") + theme_bw() +
    labs(title = "Distribution of player's value")

train = train %>% 
    mutate(value = value^(1/4)) %>% 
    rename("transformed_value" = value)

test = test %>% 
    mutate(value = value^(1/4)) %>% 
    rename("transformed_value" = value)

train %>% write_csv("..\\exploratory analysis\\train.csv")
test %>% write_csv("..\\exploratory analysis\\test.csv")

data = train

p2 = data %>% 
    ggplot(aes(x = transformed_value)) + geom_density(fill = "navy") +   
    theme_bw() + 
    labs(title = "Distribution of transformed value")

p1 + p2
```




#Checking NA's

##NA's in each observation

```{r}

data = data %>% 
    mutate(., na_count = apply(., 1, function(x) sum(is.na(x))))

data %>% 
    filter(na_count > 0) %>% 
    ggplot(aes(x = na_count)) + geom_histogram(fill = "navy") + theme_bw() +
    labs(title = "Distribution of NA's in goal-keeper data",
         x = "number of missing values") +
    theme(plot.title = element_text(size = 10))

```


```{r fig.height=13, fig.width=11}

library(VIM)
#In the following plots, red color represents missing data

matrixplot(data[,1:20],
           labels = c("age","nationality","potential","transvalue","special",
                    "accelerat","aggression","agility","balance","ball_con",
                    "composure","crossing","curve","dribbling","finishing",
                    "free_kick","gk_diving","gk_handle","gk_kick","gk_posi"))

matrixplot(data[,21:39], 
           labels = c("gk_reflex", "hand_accu", "intercept", "jumping",
                      "long_pass", "long_shots", "marking", "penalties",
                      "position", "reactions", "short pass", "shot pwer",
                      "slid_tack", "sprint", "stamina", "std_tack",
                      "strength", "vision", "volley"))

```

##NA's in each variable

```{r}
na_col = colSums(is.na(data)) %>% 
    as.list() %>% 
    as.data.frame() %>% 
    select(-na_count) %>% 
    gather(age:volleys, key = "variable", value = "num_of_na")

na_col %>% 
    filter(num_of_na > 0) %>% 
    mutate(variable = fct_reorder(variable, num_of_na)) %>% 
    ggplot(aes(x = variable, y = num_of_na)) + 
  geom_col(fill = "navy") +  
  theme(legend.position = "bottom") + 
  labs(title = "number of missing values for variables that have NA(s)", 
       y = "number of missing values") +
    coord_flip() + theme(axis.text.x = element_text(face = "plain", 
                                                    color = "black", 
                                                    size = 8)) + 
    theme_bw()
```



#Tables for descriptive statistics

```{r}
descrip_list = data %>% 
    skimr::skim_to_list()

descrip_list[[1]] %>% 
    select(variable, n_unique, missing) %>% 
    dplyr::rename("unique levels" = n_unique) %>% 
    knitr::kable(caption = "Factor variables")
    

```

```{r}
bind_rows(descrip_list[[2]] , descrip_list[[3]]) %>% 
    dplyr::select(variable, 
                Min = p0,
                `1st Q` = p25,
                Mean = mean,
                Median = p50,
                `3rd Q` = p75,
                Max = p100,
                `Std Dev` = sd,
                missing) %>% 
  knitr::kable(digits = 3, caption = "Integer/numeric variables")
```


#Figures for descriptive statistics


##Check the distribution for each numeric/integer predictor

```{r fig.height=13, fig.width=11}
#library(gridExtra)

p1 = data[,1:18] %>%
    select(-transformed_value) %>% 
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density() +
    theme_bw() 
    

p2 = data[,19:34] %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density() +
    theme_bw()

p1/p2
```

\newpage

```{r fig.height=2.6, fig.width=5}
data[,35:39] %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density()  +
    theme_bw()
```


##plot for variables

###for factor variables


* nationality

plot the situation for nations with the most number of players:

```{r}
nation_box = data %>% 
    mutate(nationality = fct_lump(nationality, 12)) %>% 
    mutate(nationality = fct_infreq(nationality)) %>% 
    mutate(nationality = fct_rev(nationality)) %>% 
    #move "Other" level to the last:
    mutate(nationality = fct_relevel(nationality, "Other", after = 0)) %>% 
    ggplot(aes(x = nationality, y = transformed_value)) +
geom_boxplot() +
theme(legend.position = "bottom") +
labs( x = NULL) +
coord_flip() + 
    theme(axis.text.x = element_text(face = "plain", 
                                     color = "black", 
                                     size = 8)) + 
    theme_bw()

nation_hist = data %>% 
    #If focus on most common nations:
    mutate(nationality = fct_lump(nationality, 12)) %>% 
    #nations that have fewer players will be denoted as "Other"
    mutate(nationality = fct_infreq(nationality)) %>% 
    mutate(nationality = fct_rev(nationality)) %>% 
    #move "Other" level to the last:
    mutate(nationality = fct_relevel(nationality, "Other", after = 0)) %>% 
    ggplot(aes(x = nationality)) +
    geom_bar(fill = "navy") +
    theme(legend.position = "bottom") +
    labs(title = "Player count/transformed_value by nationality", 
         subtitle = "Nations with most players. This plot suggests that
         players' \n transformed_values vary between different nations") +
    coord_flip() + 
    theme(axis.text.x = element_text(face = "plain", 
                                     color = "black", 
                                     size = 8))  + 
    theme_bw()

nation_hist + nation_box
```

plot the nations with highest players' transformed_values:

```{r}
nation_box = data %>%
    group_by(nationality) %>% 
    mutate(med_by_nation = median(transformed_value)) %>% 
    ungroup() %>% 
    mutate(nationality = fct_reorder(nationality, med_by_nation)) %>% 
    ggplot(aes(x = nationality, y = transformed_value)) +
    geom_boxplot() +
    theme(legend.position = "bottom") +
    labs( x = NULL) +
    coord_flip() + 
    theme(axis.text.x = element_text(face = "plain", 
                                     color = "black", 
                                     size = 8)) + 
    theme_bw()

nation_hist = data %>%
    group_by(nationality) %>% 
    mutate(med_by_nation = median(transformed_value)) %>% 
    ungroup() %>% 
    mutate(nationality = fct_reorder(nationality, med_by_nation)) %>% 
    ggplot(aes(x = nationality)) +
    geom_bar(fill = "navy") +
    theme(legend.position = "bottom") +
    labs(title = "Player count/transformed_value by nation", 
         subtitle = "Nations with highest median player transformed_values.
         Those with the highest player\n transformed_values typically have
         very little player data recorded.") +
    coord_flip() + 
    theme(axis.text.x = element_text(face = "plain", 
                                     color = "black", 
                                     size = 8))  + 
    theme_bw()

nation_hist + nation_box
```

###for int/num variables


```{r fig.height=15, fig.width=11}
# matrix of predictors

data_num  = data %>% 
    keep(is.numeric) %>% 
    select(-na_count, -transformed_value) 

#for factor variables
data_fct  = data %>%  
    select(-transformed_value) %>% 
    select_if(~ is.factor(.))


# vector of response
y <- data$transformed_value

featurePlot(data_num,
y,
plot = "scatter",
span = .5,
labels = c("Predictors","Y"),
type = c("p", "smooth"),
layout = c(4, 10))

```

\newpage



###FeaturePlots for factor variables

```{r}
featurePlot(data$transformed_value, data_fct$nationality, "box", labels = c("Nationality","Y"))
featurePlot(data$transformed_value, data_fct$club, "box", labels = c("Club","Y"))
```

##Correlation plot

```{r fig.height=12, fig.width=12}
library(corrplot)

cor_data = data %>% 
    filter(na_count == 0) %>% 
    select(-nationality, -na_count)

corrplot(cor(cor_data), tl.cex = 1.2)
```






